---
title: "part1"
author: "dogvile"
date: "Sep 25, 2015"
output: html_document
---

### Overview:
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with function rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lamda and the standard deviation is also 1/lambda. For current purposes the lambda is 0.2 for all of the simulations. 


> Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.  You should:  
> 1. Show the sample mean and compare it to the theoretical mean of the distribution.  
> 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.  
> 3. Show that the distribution is approximately normal.  

### Simulations: 
The simulation can be built in R using the `rexp(n,l)` function, where `n` is the number of exponentials per set, and `l`, or lamda is the rate (a constant that for  these purposes is 0.2).
We store the matrix with the simulated points(given num,samle,rate and type of distribution) in a variable 'simulation'

```{r, echo=FALSE}
set.seed(999564646); #Pseudo Random Number Generator

lamda<-0.2        # lambda
num<-40         # number of exponentials
sample<-1000      # number of simulations

#store in matrix 1000 simulations following exponential distribution
simulation <- matrix(rexp(num*sample, rate=lamda), num, sample)

```


### Sample Mean versus Theoretical Mean: 
The formula of theorical mean is 1/lamda where lamda in this case is 0.2. Also
we will calculate the mean of simulated observations. Based on results are quite close

```{r}
row_means <- rowMeans(simulation)
sample_mean <- mean(row_means)
theoritical_mean <- 1/lamda
```

The sample mean of observation data is 4.999 which is very close to the theoretical mean of 5.0.  When we plot the sample set below and place a vertical line on both mean results, the histograms look almost identical.

```{r, echo=FALSE}


hist(row_means, breaks=5, prob=TRUE,col="grey",
     main="Distribution of samples' mean,
     drawn from exponential distribution with lambda=0.2",
     xlab="")
# density of the averages of samples
lines(density(row_means))

# theoretical center of distribution based on formula of rate 1/lamda so value 5 as center
abline(v=1/lamda, col="red")

# theoretical density of the averages of samples
xfit <- seq(min(row_means), max(row_means), length=100)
yfit <- dnorm(xfit, mean=1/lamda, sd=(1/lamda/sqrt(sample)))
lines(xfit, yfit, pch=22, col="red", lty=2)

# add legend
legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("black", "red"))
```

### Sample Variance versus Theoretical Variance: 
We can futher compare by defining the theoretical variance as 1/lambda squared, divided by the number of exponential observations, or `num`.  

```{r}
theoretical_variance<- (1/lamda)^2/num;# theoretical variance
sample_variance <- var(row_means)      #variance of simulated observations
```



### Distribution: 
As indicator of distribution we will use q-q plot which shows a normal distribution(points across the line)
```{r}
qqnorm(row_means); qqline(row_means)

```